---

title: "Beer Reviews: Statistical Analysis"
author: "Sergey Mouzykin"
date: "November 25, 2017"
output: rmarkdown::github_document

---


```{r Libraries and Data Frame, message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggthemes)
beer_reviews <- read_csv("beer_reviews_clean.csv")
beer_reviews = as_data_frame(beer_reviews)
```

### Summary Statistics

```{r Functions, message=FALSE, warning=FALSE, include=FALSE}

reds = brewer.pal(9, 'Reds')
blues = brewer.pal(9, 'Blues')
theme_blue = theme(panel.background = element_blank(),
                  legend.key = element_rect(fill = 'gray', color = "#3182BD", size = 1),
                  legend.background = element_rect(fill = 'gray', color = "#3182BD", size = 1),
                  legend.position = 'right',
                  legend.direction = 'vertical',
                  strip.background = element_blank(),
                  plot.background = element_rect(fill = 'gray', color = "#3182BD", size = 2),
                  panel.grid = element_line(linetype = 8),
                  axis.line = element_line(color = "#08519C"),
                  axis.ticks = element_line(color = "#08519C"),
                  axis.ticks.x = element_line(color = "black", size = 1),
                  axis.ticks.y = element_line(color = "black", size = 1),
                  strip.text = element_text(size = 16, color = 'red'),
                  axis.title.y = element_text(color = 'black', hjust = 0.5, face = "italic"),
                  axis.title.x = element_text(color = 'black', hjust = 0.5, face = "italic"),
                  plot.title = element_text(color = 'black', hjust = '0.5'),
                  axis.text = element_text(color = "black"))


stats_function_2 = function(DF, col_name, summary = FALSE){
  # DF has to be a list
  if (summary == TRUE){
    summary(DF)
  }
  else if(summary == FALSE){
    DF = list(DF)  
    sd_list     = lapply(DF, sd)
    var_list    = lapply(DF, var)
    mean_list   = lapply(DF, mean)
    median_list = lapply(DF, median)
    max_list    = lapply(DF, max)
    min_list    = lapply(DF, min)
    yy          = c(sd_list, var_list, mean_list, median_list, max_list, min_list)
    rows        = c('standard deviation', 'variance', 'mean', 'median', 'max', 'min')
    columns     = c(col1 = col_name)
    sd_matrix   = matrix(yy, byrow = TRUE, nrow = length(rows), ncol = length(columns))
    colnames(sd_matrix) = columns
    rownames(sd_matrix) = rows
    sd_matrix
  }
  
}

```

### Beer Name

Summarizing the standard deviation, variance, mean, median, max, and min for every beer style.

```{r Aspects summary, echo=FALSE, message=FALSE, warning=FALSE}

# for every beer style and beer name summarize, SD, VAR, MEAN, MEDIAN, MAX, MIN
stats_function_2(beer_reviews$overall, 'Overall Rating Summary')
stats_function_2(beer_reviews$taste, 'Taste Rating Summary')
stats_function_2(beer_reviews$aroma, 'Aroma Rating Summary')
stats_function_2(beer_reviews$appearance, 'Appearance Rating Summary')
stats_function_2(beer_reviews$palate, 'Palate Rating Summary')


```


The min and max for two columns, overall and appearance, are 0 and 5 respectively. It's worth taking a closer look at those ratings since the other aspects range from 1 to 5.

Looking at the mean and standard deviation (SD) of overall column where the overall rating is given a zero, it is interesting to find that in some of the more extreme scenarios, the ratings between different aspects vary greatly. In particular, for the beer, *Latter Days Stout*, we can find that a rating of zero for the overall was given out while at the same time its aroma was rated a four. It seems contradictory to rate the overall as a zero while enjoying at least one aspect of the beer, in particular, its aroma. 

```{r Beer name stats, echo=FALSE, message=FALSE, warning=FALSE}

# the min and max for two columns, overall and appearance, are 0 and 5 respectively. It's worth taking a closer look at those ratings
# look at the mean and SD of overall column where the overall rating is given a zero
beer_reviews_dist = 
  beer_reviews %>%
  group_by(beer_name, overall, taste, aroma, appearance, palate) %>%
  summarise(
    beer_name_count = n(), 
    overall_mean = mean(overall), overall_sd = sd(overall)) %>%
  arrange(overall) %>%
  filter(overall == 0) %>%
  print(n = 10)


# tally each beer name and calculate SD for the 5 observations
beer_reviews_SD = 
  beer_reviews %>%
  group_by(beer_name, general_beer_style) %>%
  summarise(beer_name_count = n(),
            
            overall_sd    = sd(overall), 
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(overall_sd) %>%
  na.omit %>%
  print(n = 20)

stats_function_2(beer_reviews_SD$beer_name_count, 'Beer Review Count')

```

# Plot the Standard Deviation Distribution

In support to the histogram, we can also calculate the mean, median, standard deviation, maximum rating, and minimum rating. Essentially, the deviation can tell us how divisive people are among the ratings. Generally, it appears that people will agree on a rating value to within about 3/4 of a point. However, the deviation will vary with the beer name. 

```{r}
# PLOT
# overall_sd histogram
ggplot(beer_reviews_SD, aes(x = overall_sd)) + 
  geom_histogram(binwidth = 0.1, fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Overall Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_blue
stats_function_2(beer_reviews_SD$overall_sd, 'Overall SD Summary')

ggplot(beer_reviews_SD, aes(x = taste_sd)) + 
  geom_histogram(binwidth = 0.1, fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Taste Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_blue
stats_function_2(beer_reviews_SD$taste_sd, 'Taste SD Summary')

ggplot(beer_reviews_SD, aes(x = aroma_sd)) + 
  geom_histogram(binwidth = 0.1, fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Aroma Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_blue
stats_function_2(beer_reviews_SD$aroma_sd, 'Aroma SD Summary')

ggplot(beer_reviews_SD, aes(x = appearance_sd)) + 
  geom_histogram(binwidth = 0.1, fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Appearance Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_blue
stats_function_2(beer_reviews_SD$appearance_sd, 'Appearance SD Summary')
  
ggplot(beer_reviews_SD, aes(x = palate_sd)) + 
  geom_histogram(binwidth = 0.1, fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Palate Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_blue 
stats_function_2(beer_reviews_SD$palate_sd, 'Palate SD Summary')

# NOTE: 
# with only 1 or 2 reviews, disagreements vary greatly among the 5 observations
# 1. lower deviation implies agreement between different reviews
# 2. higher deviation implies disagreement between reviewers

# GRAPH NOTE:
# The outliers contained in the graph are due to low amount of reviews and the disagreements between reviewers

# FIND the least amount of disagreements by filtering the SD for each observation. Smaller deviation = less disagreement
# filter(sd between(0, 0.25), name_count >=5)
beer_reviews_sd_filter = 
  beer_reviews %>%
  group_by(beer_name, general_beer_style) %>%
  summarise(beer_name_count = n(), 
            overall_sd    = sd(overall), 
            taste_sd      = sd(taste), 
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  filter(beer_name_count >= 5, 
         between(overall_sd, 0, 0.30), 
         between(taste_sd, 0, 0.30), 
         between(aroma_sd, 0, 0.30), 
         between(appearance_sd, 0, 0.30), 
         between(palate_sd, 0, 0.30)) %>%
  arrange(overall_sd, taste_sd, aroma_sd, appearance_sd, palate_sd) %>%
  print(n = 20)


```


### General Beer Style

```{r General beer style, echo=FALSE, message=FALSE, warning=FALSE}
# tally each General_beer style and find SD
beer_reviews_style_sd = 
  beer_reviews %>%
  group_by(general_beer_style) %>%
  summarise(style_count = n(),
            
            overall_sd    = sd(overall), 
            overall_mean  = mean(overall),
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(overall_mean) %>%
  na.omit %>%
  print(n = 25)

# overall_mean shows the most disliked beer style: low alcohol
stats_function_2(beer_reviews_style_sd$style_count, 'Style Count')
stats_function_2(beer_reviews_style_sd$overall_sd, 'Overall SD')
stats_function_2(beer_reviews_style_sd$overall_mean, 'Overall Mean')
stats_function_2(beer_reviews_style_sd$aroma_sd, 'Aroma SD')
stats_function_2(beer_reviews_style_sd$taste_sd, 'Taste SD')
stats_function_2(beer_reviews_style_sd$appearance_sd, 'Appearance SD')
stats_function_2(beer_reviews_style_sd$palate_sd, 'Palate SD')

```

We can find the least disliked beer style by looking at the mean of the overall (shown above). However, the standard deviation (1.00) is quite high in this case which implies disagreement and the ratings can vary by one point. Therefore, we cannot absolutely conclude that the low alcohol beer is rated the worst. Although, we can say that it is one the least appreaciated beer styles among others.

Also, we can find the beer styles which are relatively best rated (shown below). 

### Beer Style

```{r Beer style, echo=FALSE, message=FALSE, warning=FALSE}

# tally each beer_style and find SD
beer_reviews_style_sd2 = 
  beer_reviews %>%
  group_by(beer_style) %>%
  summarise(style_count = n(),
            
            overall_sd    = sd(overall), 
            overall_mean  = mean(overall),
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(desc(overall_mean)) %>%
  na.omit %>%
  print(n = 50)

# missing values are created because there are too few amount of reviews and therefore the SD cannot be calculated
# missing values need to be omitted as a result
stats_function_2(beer_reviews_style_sd2$overall_sd, 'Overall SD')
stats_function_2(beer_reviews_style_sd2$overall_mean, 'Overall Mean')
stats_function_2(beer_reviews_style_sd2$taste_sd, 'Taste SD')
stats_function_2(beer_reviews_style_sd2$aroma_sd, 'Aroma SD')
stats_function_2(beer_reviews_style_sd2$appearance_sd, 'Appearance SD')
stats_function_2(beer_reviews_style_sd2$palate_sd, 'Palate SD')

```


### Beer ABV (alcohol by volume)

This plot shows us that the ABV is actually a right-skewed distribution due to some beers having a a very high alcohol by volume content. The minimum and maximum of ABV content are 0.01% and 57.7%, respectively. The mean lies at 7.04, the median at 6.6, and the standard deviation is 2.27. We can also infer that the distribution is right-skewed due to the median being lower than the mean. Although the values range from 0.01 to 57.7, in theory, about 95% of these values should lie within two standard deviations from the mean.

```{r}
# reduce the amount of data points with the sample_n()
# this will produce the plots much faster than plotting all 1.5 million reviews
beer_reviews_10k = sample_n(beer_reviews, size = 10000)

# PLOT: Beer ABV distribution
ggplot(beer_reviews_10k, aes(x = beer_abv)) + 
  geom_histogram(binwidth = 0.5, position = 'dodge', fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Alcohol by Volume (%)', breaks = seq(0, 60, by = 2)) +
  scale_y_continuous('Beer Count') +
  ggtitle('Beer ABV Distribution') +
  theme_blue

stats_function_2(beer_reviews$beer_abv, 'Beer ABV Summary')

# HISTOGRAM
ggplot(beer_reviews_10k, aes(x = as.factor(overall), fill = general_beer_style)) +
  geom_histogram(binwidth = 15, stat = 'count') +
  #scale_x_continuous(limits = c(1,20)) + 
  theme_classic()

# ABV factor and beer style
ggplot(beer_reviews_10k, aes(x = beer_abv_factor, fill = general_beer_style)) + 
  geom_histogram(stat = 'count', position = 'dodge') +
  scale_x_discrete('Beer ABV') + 
  theme_blue


```




```{r}

# --- Point Plot
# overall vs. beer ABV, colored by beer style
ggplot(beer_reviews_10k, aes(x = overall, y = beer_abv, col = general_beer_style)) + 
  geom_point(alpha = 0.3) +
  geom_jitter() +
  theme_blue

# Overall vs. Style, colored by beer style
ggplot(beer_reviews_10k, aes(x = overall, y = general_beer_style, col = beer_abv_factor)) + 
  geom_point(alpha = 0.1, shape = 21) +
  geom_jitter() + 
  theme_blue

ggplot(beer_reviews_10k, aes(x = beer_abv, y = general_beer_style, col = as.factor(overall))) + 
  geom_point(alpha = 0.1, shape = 21) +
  geom_jitter() + 
  theme_blue

# --- Histograms
# Overall and Beer ABV distribution
ggplot(beer_reviews_10k, aes(x = overall, fill = beer_abv_factor)) + 
  geom_histogram(binwidth = 0.25) +
  theme_blue
# Overall and Beer Styles
ggplot(beer_reviews_10k, aes(x = overall, fill = general_beer_style)) + 
  geom_histogram(binwidth = 0.25) +
  theme_blue

```

The graphs above helps visualize the overall rating distribution amongst the beer styles and their respective alcohol content.


```{r Five Aspects, echo=FALSE, message=FALSE, warning=FALSE}
# palate vs overall
ggplot(beer_reviews_10k, aes(x = palate, y = overall)) + 
  geom_point(alpha = 0.6) +
  stat_smooth(method = 'loess', se = TRUE) + 
  geom_jitter() +
  theme_blue

# taste vs overall
ggplot(beer_reviews_10k, aes(x = taste, y = overall)) + 
  geom_point(alpha = 0.6) +
  stat_smooth(method = 'loess', se = TRUE) + 
  geom_jitter() +
  theme_blue

# aroma vs overall
ggplot(beer_reviews_10k, aes(x = aroma, y = overall)) + 
  geom_point(alpha = 0.6) +
  stat_smooth(method = 'loess', se = TRUE) + 
  geom_jitter() +
  theme_blue

# appearance vs overall
ggplot(beer_reviews_10k, aes(x = appearance, y = overall)) + 
  geom_point(alpha = 0.6) +
  stat_smooth(method = 'loess', se = TRUE) + 
  geom_jitter() +
  theme_blue
cor
```

Plotting overall against other aspects, we can observe a linear correlation emerging.

# Plot: Distribution of Ratings 

These plots below help visualize the rating distribution of the five aspects, overall, taste, aroma, appearance, and palate.

```{r}

# ------ Histogram Plot -------
# Overall Rating Distribution
ggplot(beer_reviews_10k, aes(x = overall)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(beer_reviews_10k, aes(x = taste)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(beer_reviews_10k, aes(x = aroma)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(beer_reviews_10k, aes(x = appearance)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(beer_reviews_10k, aes(x = palate)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

```




### Profile Names

Here we can find the summary statistics for the profile names columns. 

Amount of profile names: 33,388

Out of these names, 19,821 of them submitted five or less reviews.

```{r}
# number of profile names
beer_reviews %>% 
  select(profile_name) %>% 
  distinct()

# amount of reviews by each person
beer_reviews %>% 
  group_by(profile_name) %>% 
  tally(sort = TRUE)

# find total amount of profile names
profile_names = 
  beer_reviews %>% 
  group_by(profile_name) %>%
  summarise(reviewed_amount = n(),
            overall_mean = mean(overall),
            overall_sd = sd(overall)) %>%
  arrange(desc(reviewed_amount)) %>%
  filter(reviewed_amount >= 3) %>%
  
  print(n = 20)
profile_names # 33,388 total profile names
glimpse(profile_names)


# find lowest activity by filtering review amount 
beer_reviews %>% 
  group_by(profile_name) %>%
  summarise(review_amount = n()) %>%
  arrange(desc(review_amount)) %>%
  filter(between(review_amount, 1, 5))
# 19,821 names submitted 5 or less reviews


```

For each profile name, we can summarize the overall ratings that were given out as well as the amount of reviews by each person.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

stats_function_2(profile_names$reviewed_amount, 'Review Amount')
stats_function_2(profile_names$overall_mean, 'Profile Name')

```


