---
title: "Statistical Analysis"
author: "Sergey Mouzykin"
date: "November 19, 2017"
output: rmarkdown::github_document
---

At this time, you should have a data set for your capstone project that you've cleaned up and wrangled to get into shape that's ready for analysis. Here are some suggested steps that might get you going.
Get going by asking the following questions and looking for the answers with some code and plots:

Can you count something interesting?

Can you find some trends (high, low, increase, decrease, anomalies)?

Can you make a bar plot or a histogram?

Can you compare two related quantities?

Can you make a scatterplot?

Can you make a time-series plot?

Having made these plots, what are some insights you get from them? Do you see any correlations? Is there a hypothesis you would like to investigate further? What other questions do they lead you to ask?

```{r Libraries and Data Frame, message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(tibble)
library(RColorBrewer)
library(ggthemes)
library(GGally)
beer_reviews <- read_csv("beer_reviews_clean.csv")
beer_reviews = as_data_frame(beer_reviews)
```

### Summary Statistics


```{r Functions, message=FALSE, warning=FALSE, include=FALSE}
# ------ SUMMARY STATISTICS ------

stats_function_2 = function(DF, col_name, summary = FALSE){
  # DF has to be a list
  if (summary == TRUE){
    summary(DF)
  }
  else if(summary == FALSE){
    DF = list(DF)  
    sd_list     = lapply(DF, sd)
    var_list    = lapply(DF, var)
    mean_list   = lapply(DF, mean)
    median_list = lapply(DF, median)
    max_list    = lapply(DF, max)
    min_list    = lapply(DF, min)
    yy          = c(sd_list, var_list, mean_list, median_list, max_list, min_list)
    rows        = c('standard deviation', 'variance', 'mean', 'median', 'max', 'min')
    columns     = c(col1 = col_name)
    sd_matrix   = matrix(yy, byrow = TRUE, nrow = length(rows), ncol = length(columns))
    colnames(sd_matrix) = columns
    rownames(sd_matrix) = rows
    sd_matrix
  }
  
}

```

### Beer Name

```{r Beer name stats, echo=FALSE, message=FALSE, warning=FALSE}
# ---- BEER NAME: Summary Stats ----

# for every beer style and beer name summarize, SD, VAR, MEAN, MEDIAN, MAX, MIN
stats_function_2(beer_reviews$overall, 'Overall Rating Summary')
stats_function_2(beer_reviews$taste, 'Taste Rating Summary')
stats_function_2(beer_reviews$aroma, 'Aroma Rating Summary')
stats_function_2(beer_reviews$appearance, 'Appearance Rating Summary')
stats_function_2(beer_reviews$palate, 'Palate Rating Summary')

# the min and max for two columns, overall and appearance, are 0 and 5 respectively. It's worth taking a closer look at those ratings
# look at the mean and SD of overall column where the overall rating is given a zero
beer_reviews_dist = 
  beer_reviews %>%
  group_by(beer_name, overall, taste, aroma, appearance, palate) %>%
  summarise(
    beer_name_count = n(), 
    overall_mean = mean(overall), overall_sd = sd(overall)) %>%
  arrange(overall) %>%
  filter(overall == 0) %>%
  print(n = 10)


# tally each beer name and calculate SD for the 5 observations
beer_reviews_SD = 
  beer_reviews %>%
  group_by(beer_name, general_beer_style) %>%
  summarise(beer_name_count = n(),
            
            overall_sd    = sd(overall), 
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(overall_sd) %>%
  na.omit %>%
  print(n = 20)
glimpse(beer_reviews_SD)


stats_function_2(beer_reviews_SD$beer_name_count, 'Beer Review Count')

# plot distribution of SD
overall_sd_plot = 
  ggplot(beer_reviews_SD, aes(x = overall_sd)) + 
  geom_histogram(binwidth = 0.05, position = 'dodge', fill = 'gray', col = 'black') +
  theme_classic()
stats_function_2(beer_reviews_SD$overall_sd, 'Overall SD Summary')

taste_sd_plot = 
  ggplot(beer_reviews_SD, aes(x = taste_sd)) + 
  geom_histogram(binwidth = 0.05, position = 'dodge', fill = 'gray', col = 'black') +
  theme_classic()
stats_function_2(beer_reviews_SD$taste_sd, 'Overall SD Summary')

aroma_sd_plot = 
ggplot(beer_reviews_SD, aes(x = aroma_sd)) + 
  geom_histogram(binwidth = 0.05, position = 'dodge', fill = 'gray', col = 'black') +
  theme_classic()
stats_function_2(beer_reviews_SD$aroma_sd, 'Overall SD Summary')

appearance_sd_plot = 
  ggplot(beer_reviews_SD, aes(x = appearance_sd)) + 
  geom_histogram(binwidth = 0.05, position = 'dodge', fill = 'gray', col = 'black') +
  theme_classic()
stats_function_2(beer_reviews_SD$appearance_sd, 'Overall SD Summary')

palate_sd_plot = 
  ggplot(beer_reviews_SD, aes(x = palate_sd)) + 
  geom_histogram(binwidth = 0.05, position = 'dodge', fill = 'gray', col = 'black') +
  theme_classic()
stats_function_2(beer_reviews_SD$palate_sd, 'Overall SD Summary')

```

### General Beer Style

```{r General beer style, echo=FALSE, message=FALSE, warning=FALSE}
# ---- GENERAL BEER STYLE: Summary Stats ----

# tally each General_beer style and find SD
beer_reviews_style_sd = 
  beer_reviews %>%
  group_by(general_beer_style) %>%
  summarise(style_count = n(),
            
            overall_sd    = sd(overall), 
            overall_mean  = mean(overall),
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(overall_mean) %>%
  na.omit %>%
  print(n = 25)

# overall_mean shows the most disliked beer style: low alcohol

stats_function_2(beer_reviews_style_sd$style_count, 'Style Count')
stats_function_2(beer_reviews_style_sd$overall_sd, 'Overall SD')
stats_function_2(beer_reviews_style_sd$overall_mean, 'Overall Mean')
stats_function_2(beer_reviews_style_sd$aroma_sd, 'Aroma SD')
stats_function_2(beer_reviews_style_sd$taste_sd, 'Taste SD')
stats_function_2(beer_reviews_style_sd$appearance_sd, 'Appearance SD')
stats_function_2(beer_reviews_style_sd$palate_sd, 'Palate SD')

```

### Beer Style

```{r Beer style, echo=FALSE, message=FALSE, warning=FALSE}
# ---- BEER STYLE: Summary Stats ----

# tally each beer_style and find SD
beer_reviews_style_sd2 = 
  beer_reviews %>%
  group_by(beer_style) %>%
  summarise(style_count = n(),
            
            overall_sd    = sd(overall), 
            overall_mean  = mean(overall),
            taste_sd      = sd(taste),
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  
  #filter(beer_name_count >= 5) %>%
  arrange(desc(overall_mean)) %>%
  na.omit %>%
  print(n = 50)

# missing values are created because there are too few amount of reviews and therefore the SD cannot be calculated
# missing values need to be omitted as a result
stats_function_2(beer_reviews_style_sd2$overall_sd, 'Overall SD')
stats_function_2(beer_reviews_style_sd2$overall_mean, 'Overall Mean')
stats_function_2(beer_reviews_style_sd2$taste_sd, 'Taste SD')
stats_function_2(beer_reviews_style_sd2$aroma_sd, 'Aroma SD')
stats_function_2(beer_reviews_style_sd2$appearance_sd, 'Appearance SD')
stats_function_2(beer_reviews_style_sd2$palate_sd, 'Palate SD')

```




```{r}

# PLOT
# overall_sd histogram
ggplot(beer_reviews_SD, aes(x = overall_sd)) + 
  geom_histogram(binwidth = 0.1, col = 'black') + 
  scale_x_continuous('Overall Deviation', breaks = seq(0, 3.0, by = 0.2)) +
  theme_classic()

ggplot(beer_reviews_SD, aes(x = taste_sd)) + 
  geom_histogram(binwidth = 0.1, col = 'black') + 
  scale_x_continuous('Taste Deviation', breaks = seq(0, 3.0, by = 0.2))
  theme_classic()
  
  

# NOTE: 
# with only 1 or 2 reviews, disagreements vary greatly among the 5 observations
# 1. lower deviation implies agreement between different reviews
# 2. higher deviation implies disagreement between reviewers

# GRAPH NOTE:
# The outliers contained in the graph are due to low amount of reviews and the disagreements between reviewers

# FIND the least amount of disagreements by filtering the SD for each observation. Smaller deviation = less disagreement
# filter(sd between(0, 0.25), name_count >=5)
beer_reviews_sd_filter = 
  beer_reviews %>%
  group_by(beer_name, general_beer_style) %>%
  summarise(beer_name_count = n(), 
            overall_sd    = sd(overall), 
            taste_sd      = sd(taste), 
            aroma_sd      = sd(aroma), 
            appearance_sd = sd(appearance), 
            palate_sd     = sd(palate)) %>%
  filter(beer_name_count >= 5, 
         between(overall_sd, 0, 0.30), 
         between(taste_sd, 0, 0.30), 
         between(aroma_sd, 0, 0.30), 
         between(appearance_sd, 0, 0.30), 
         between(palate_sd, 0, 0.30)) %>%
  arrange(overall_sd, taste_sd, aroma_sd, appearance_sd, palate_sd) %>%
  print(n = 20)


```


```{r}
# ---- PLOTS: themes ----
# use gsub on beer_styles column and create a new column for a general beer style
beer_reviews_5k = sample_n(beer_reviews, 5000)
glimpse(beer_reviews)
theme
theme_pink <- theme(panel.background = element_blank(),
                    legend.key = element_blank(),
                    legend.background = element_blank(),
                    strip.background = element_blank(),
                    plot.background = element_rect(fill = 'pink', color = "black", size = 3),
                    panel.grid = element_blank(),
                    axis.line = element_line(color = "black"),
                    axis.ticks = element_line(color = "black"),
                    strip.text = element_text(size = 16, color = 'red'),
                    axis.title.y = element_text(color = 'red', hjust = 0.5, face = "italic"),
                    axis.title.x = element_text(color = 'red', hjust = 0.5, face = "italic"),
                    plot.title = element_text(color = 'red', hjust = '0.5'),
                    
                    axis.text = element_text(color = "black"),
                    legend.position = "none")
?theme
reds = brewer.pal(9, 'Reds')
blues = brewer.pal(9, 'Blues')
theme_blue = theme(panel.background = element_blank(),
                  legend.key = element_rect(fill = 'gray', color = "#3182BD", size = 1),
                  legend.background = element_rect(fill = 'gray', color = "#3182BD", size = 1),
                  legend.position = 'right',
                  legend.direction = 'vertical',
                  strip.background = element_blank(),
                  plot.background = element_rect(fill = 'gray', color = "#3182BD", size = 2),
                  panel.grid = element_line(linetype = 8),
                  axis.line = element_line(color = "#08519C"),
                  axis.ticks = element_line(color = "#08519C"),
                  axis.ticks.x = element_line(color = "black", size = 1),
                  axis.ticks.y = element_line(color = "black", size = 1),
                  strip.text = element_text(size = 16, color = 'red'),
                  axis.title.y = element_text(color = 'black', hjust = 0.5, face = "italic"),
                  axis.title.x = element_text(color = 'black', hjust = 0.5, face = "italic"),
                  plot.title = element_text(color = 'black', hjust = '0.5'),
                  axis.text = element_text(color = "black"))


```



```{r}
# ------ PLOTTING ------

# PLOT: Beer ABV distribution
beer_reviews_10k = sample_n(beer_reviews, size = 10000)

ggplot(beer_reviews_10k, aes(x = beer_abv)) + 
  geom_histogram(binwidth = 0.5, position = 'dodge', fill = '#3182BD', col = 'black') + 
  scale_x_continuous('Alcohol by Volume (%)', breaks = seq(0, 60, by = 2)) +
  scale_y_continuous('Beer Count') +
  ggtitle('Beer ABV Distribution') +
  theme_blue

stats_function_2(beer_reviews$beer_abv, 'Beer ABV Summary')

# PLOT: ABV and Style
ggplot(beer_reviews_10k, aes(x = general_beer_style , y = beer_abv )) + 
  geom_point() + 
  geom_jitter()

ggplot(beer_reviews_10k, aes(x = general_beer_style , y = beer_abv, fill = overall )) + 
  geom_bar(stat = 'identity')


# HISTOGRAM
ggplot(beer_reviews_10k, aes(x = beer_abv, fill = general_beer_style)) + 
  geom_histogram(binwidth = 10, position = 'dodge', alpha = 0.4) 
  #scale_x_continuous( breaks = seq(0, 60, by = 2)) +
  #theme_blue

ggplot(beer_reviews_5k, aes(x = as.factor(overall), fill = general_beer_style)) + # ??????????????????
  geom_histogram(binwidth = 15, stat = 'count') +
  #scale_x_continuous(limits = c(1,20)) + 
  theme_classic()

# ABV factor and beer style
ggplot(beer_reviews_5k, aes(x = beer_abv_factor, fill = general_beer_style)) + 
  geom_histogram(stat = 'count', position = 'dodge') +
  scale_x_discrete('Beer ABV') + 
  theme_blue


# BAR GRAPH
# beer style vs overall
ggplot(beer_reviews_5k, aes(x = general_beer_style, y = overall, fill = general_beer_style)) + 
  geom_bar(stat = 'identity') +
  scale_x_discrete('Beer Style',  expand = c(0.1, 0)) 
  #scale_y_discrete('Overall Rating', breaks = seq(0, 5, by = 1)) +
  #theme_classic()

```


```{r}

# reduce the amount of data points with the sample_n()
reviews_10k = sample_n(beer_reviews, size = 10000)
glimpse(reviews_10k)

# --- Point Plot
# overall vs. beer ABV, colored by beer style
ggplot(reviews_10k, aes(x = overall, y = beer_abv, col = general_beer_style)) + 
  geom_point(alpha = 0.3) +
  geom_jitter() +
  theme_blue

# filter out Lager in particular with its respective ABV, and ratings
# plot Lager overall ratings vs. the ABV

# --- Point Plot
# Overall vs. Style, colored by beer style
ggplot(reviews_10k, aes(x = overall, y = general_beer_style, col = beer_abv_factor)) + 
  geom_point(alpha = 0.1, shape = 21) +
  geom_jitter() + 
  theme_blue
# Overall and Beer ABV distribution
ggplot(reviews_10k, aes(x = overall, fill = beer_abv_factor)) + 
  geom_histogram(binwidth = 0.25) +
  theme_blue
# Overall and Beer Styles
ggplot(reviews_10k, aes(x = overall, fill = general_beer_style)) + 
  geom_histogram(binwidth = 0.25) +
  theme_blue

ggplot(reviews_10k, aes(x = beer_abv, y = general_beer_style, col = as.factor(overall))) + 
  geom_point(alpha = 0.1, shape = 21) +
  geom_jitter() + 
  theme_blue

# ------ Histogram Plot -------
# Overall Rating Distribution
ggplot(reviews_10k, aes(x = overall)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(reviews_10k, aes(x = taste)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(reviews_10k, aes(x = aroma)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(reviews_10k, aes(x = appearance)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue

ggplot(reviews_10k, aes(x = palate)) + 
  geom_histogram(binwidth = 0.25, position = 'dodge', fill = '#3182BD', col = 'black') +
  theme_blue


```




```{r}
# POINT PLOTS
# overall vs. ABV
ggplot(beer_reviews_5k, aes(x = overall, y = beer_abv, col = general_beer_style)) + 
  #geom_point(alpha = 0.3) +
  stat_smooth(method = 'loess', se = FALSE) +
  #geom_jitter() +
  theme_blue
  
# overall vs. beer style
ggplot(beer_reviews_5k, aes(x = overall, y = general_beer_style, col = beer_abv_factor)) + 
  geom_point(alpha = 0.3) +
  #stat_smooth(method = 'loess', se = FALSE) +
  geom_jitter() 
  #theme_blue
  
ggplot(beer_reviews_5k, aes(x = overall, y = general_beer_style, col = beer_abv_factor)) + 
  geom_point(shape = 21, alpha = 0.6) + 
  geom_jitter()  +
  #stat_smooth(method = 'loess', se = FALSE)
  theme_blue

# palate vs overall
ggplot(beer_reviews_5k, aes(x = palate, y = overall)) + 
  geom_point(alpha = 0.3) +
  stat_smooth(method = 'loess', se = TRUE) + 
  geom_jitter() +
  theme_classic()
```



```{r}
# ------ PROFILE NAMES: Summary Stats ------
# number of profile names
beer_reviews %>% 
  select(profile_name) %>% 
  distinct()

# amount of reviews by each person
beer_reviews %>% 
  group_by(profile_name) %>% 
  tally(sort = TRUE)

# find total amount of profile names
profile_names = 
  beer_reviews %>% 
  group_by(profile_name) %>%
  summarise(reviewed_amount = n(),
            overall_mean = mean(overall),
            overall_sd = sd(overall)) %>%
  arrange(desc(reviewed_amount)) %>%
  filter(reviewed_amount >= 3) %>%
  
  print(n = 20)
profile_names # 33,388 total profile names
glimpse(profile_names)

stats_function_2(profile_names$reviewed_amount, 'Profile Name')
stats_function_2(profile_names$overall_mean, 'Profile Name')

beer_reviews %>%
  select(profile_name) %>%
  contains('northyorksammy')

name1 = contains('northyorksammy', vars = beer_reviews$profile_name)
length(name1)

# find lowest activity by filtering review amount 
beer_reviews %>% 
  group_by(profile_name) %>%
  summarise(review_amount = n()) %>%
  arrange(desc(review_amount)) %>%
  filter(between(review_amount, 1, 5))
# 19,821 names submitted 5 or less reviews


```


